Domestic cooking and likelihood

### 2020-04-05 inst/cook.Rmd
use pwa() instead of working by hand

```{r,beforetraining}
H <- hyper2(pnames=c(paste("p",1:5,sep=""),"t"))  # 't' the strength of training
# observations before training:
Hbefore <- H +
    rank_likelihood(c(5,2,3,4,1)) + #  5 is the best, 4 the worst.
    rank_likelihood(c(5,3,2,4,1)) + #  5 is the best, 1 the worst.
    rank_likelihood(c(4,5,3,2,1)) + #  4 is the best, 1 the worst.
    rank_likelihood(c(5,3,4,1,2)) + #  5 is the best, 2 the worst.
    rank_likelihood(c(2,5,3,1,4))   #  2 is the best, 4 the worst.
Hbefore
```

@@

We can translate these into package idiom by defining likelihood
functions `H1after` and `H2after` for the two observations:

```{r,aftertraining1}
H1after <- H
H1after[c("p1","t")] %<>% inc()
H1after[  "p2"     ] %<>% inc()
H1after[  "p3"     ] %<>% inc()
H1after[  "p4"     ] %<>% inc()
H1after[  "p5"     ] %<>% inc()
H1after[c("p1","t","p2","p3","p4","p5")] %<>% dec()
H1after[c(         "p2","p3","p4","p5")] %<>% dec()
H1after[c(              "p3","p4","p5")] %<>% dec()
H1after[c(                   "p4","p5")] %<>% dec()
H1after[c(                        "p5")] %<>% dec()
H1after
```

```{r,aftertraining2}
H2after <- H
H2after[  "p3"     ] %<>% inc()
H2after[c("p1","t")] %<>% inc()
H2after[  "p5"     ] %<>% inc()
H2after[  "p2"     ] %<>% inc()
H2after[  "p4"     ] %<>% inc()
H2after[c("p1","t","p2","p3","p4","p5")] %<>% dec()
H2after[c("p1","t","p2",     "p4","p5")] %<>% dec()
H2after[c(         "p2",     "p4","p5")] %<>% dec()
H2after[c(         "p2",     "p4"     )] %<>% dec()
H2after[c(                   "p4"     )] %<>% dec()
H2after
```

Combine:

```{r,combineboth}
(Htot <- Hbefore + H1after + H2after)  # observations are independent!
(ML_unconstrained <- maxp(Htot))
```

We see the training entity has an estimated strength of about 0.63,
higher than all real players combined.  At the evaluate the
loglikelihood is

```{r,loglikateval}
loglik(indep(ML_unconstrained),Htot)
```

We can test the hypothesis that the training was ineffective by
constraining the optimization to $t=0$:

```{r}
small <- 1e-4
(ML_constrained <- maxp(Htot,startp=rep(0.2*(1-small*2.1),5),fcm=c(1,1,1,1,1),fcv=1-2*small))
loglik(indep(ML_constrained),Htot)
```

See how the constraint is active: the strength of the training effect
is almost exactly equal to its maximal value of `small` [as a
consistency check we can perform the same calculation with a slightly
different value for `small`:

```{r}
small <- 2e-4  # cf 1e-4 previously
ML_constrained <- maxp(Htot,startp=rep(0.2*(1-small*2.1),5),fcm=c(1,1,1,1,1),fcv=1-2*small)
loglik(indep(ML_constrained),Htot) # cf -29.973 previously
```

and we can see that this is almost the same].  Thus, the support for
our alternative hypothesis would be

```{r}
loglik(indep(ML_unconstrained),Htot)-loglik(indep(ML_constrained),Htot)
```

thus exceeding Edwards's two units of support.  We have strong
evidence to suggest that the training was effective.

>>>>

```{r,beforetraining}
H <- hyper2(pnames=c(paste("p",1:5,sep="")))  # 'S' the strength of training
# observations before training:
Hbefore <- H +
    rank_likelihood(c(5,2,3,4,1)) + #  5 is the best, 4 the worst.
    rank_likelihood(c(5,3,2,4,1)) + #  5 is the best, 1 the worst.
    rank_likelihood(c(4,5,3,2,1)) + #  4 is the best, 1 the worst.
    rank_likelihood(c(5,3,4,1,2)) + #  5 is the best, 2 the worst.
    rank_likelihood(c(2,5,3,1,4))   #  2 is the best, 4 the worst.
Hbefore
```

@@

We can translate these into package idiom with
function `pwa()` ("person with advantage"):

```{r}
Hafter <- rank_likelihood(1:5)+rank_likelihood(c(3,1,5,2,4))
pnames(Hafter) <- pnames(Hbefore)
Hafter %<>%  pwa("p1")
Hafter
```

In the above, the additional strength conferred by the training is
represented by `S`.  We can estimate the effect of training:

```{r}
pnames(Hbefore) <- c(pnames(Hbefore),"S")
maxp(Hbefore + Hafter)
```

The above suggesting that the training had a sizable effect; and we
may test the null that the training had no effect:

```{r}
specificp.gt.test(Hbefore + Hafter,"S",0)
```

thus we have strong evidence for the training being effective.
